---
title: "Project"
author: "Moustafa Abdelaziz, Erik Wong, Luke Hong"
date: "4/7/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Libraries Used}
options(warn=-1)
library(rvest)
library(xml2)
library(magrittr)
library(dplyr)
library(stringr)
require(distr)
library(RSQLite)
library(tidyverse)
library(caret)

```



Function that creates income statement by ticker by scraping data from Nasdaq.
```{r Creates an Income Statement By Ticker}
income_statement_by_ticker <- function(ticker) {
  
  #page used to scrape data regarding income statement values
  page <- read_html(paste(paste("https://www.nasdaq.com/symbol/", ticker, sep=""), "/financials", sep=""))
  income_statement <- html_text(html_nodes(page, "td+ td , .td_genTable+ td , th"))
  
  #
  income_df <- data.frame(income_statement)
  income_df <- data.frame(income_df[2:nrow(income_df),])
  names(income_df) <- c("Values")
  income_df <- income_df[income_df != ""]
  income_df <- income_df[income_df != "Operating Expenses"]
  
  company_financials <- data.frame(ticker = double(),
                                   "2018" = double(),
                                   "2017" = double(),
                                   "2016" = double(),
                                   "2015"= double())
  
  if (length(income_df) > 2) {
    for (val in seq(from=6, to=length(income_df), by=5)) {
      row <- c(as.character(income_df[val]), income_df[val+1], income_df[val+2], income_df[val+3], income_df[val+4])
      company_financials[nrow(company_financials)+1,] <- row }
    
    colnames(company_financials)[1] <- ticker
  }
  company_financials
}

```



The three compiled financial statements of all S&P 500 companies were scraped from Nasdaq and stored in a csv for easy access.
```{r Read Three Compiled Financial Statements of All S&P 500 Companies }

final_income <- read.csv("compiled_income_statements.csv", sep=",", TRUE)
final_balance <- read.csv("compiled_balance_sheets.csv", sep=",", TRUE)
final_cash <- read.csv("compiled_cash_flows.csv", sep=",", TRUE)

```


Functions used for cleaning our data from the compiled financial statements.
```{r cleaning functions}
clean_value <- function(x) {
  
  if (!is.na(x)) {
    if (substr(x, nchar(x), nchar(x)) == "B") {
      x <- as.numeric(substr(x, 0, nchar(x) -1)) * 1000000000
    }
    
    else if (substr(x, nchar(x), nchar(x)) == "M") {
      x <- as.numeric(substr(x, 0, nchar(x) -1)) * 1000000
    }
    
    else if (substr(x, nchar(x), nchar(x)) == "k") {
      x <- as.numeric(substr(x, 0, nchar(x) -1)) * 1000
    }
    
    else if (substr(x, nchar(x), nchar(x)) == "%") {
      x <- as.numeric(substr(x, 0, nchar(x) -1)) / 100
    }
  }
  x
}

clean_number <- function(x) {
  x <- gsub("[,]", "", x)
  x <- gsub("[(]", "-", x)
  x <- gsub("[)]", "", x)
  x <- gsub("[$]", "", x)
  x <- as.double(x)
  x
}

same_names <- function(df1, df2) {
  
  boolean <- TRUE
  
  for (name1 in names(df1)) {
    if (!(name1 %in% names(df2))) {
      boolean <- FALSE
    }
  }
  
  boolean
}


```


Compiles all Tickers from the S&P 500 and their relative industries.
```{r Tickers and their relative Industries in the S&P 500 }
ticker_url <- read_html("https://en.wikipedia.org/wiki/List_of_S%26P_500_companies")
ticker_and_industry_list <- html_text(html_nodes(ticker_url, "#constituents td:nth-child(2) , #constituents td:nth-child(4)"))

ticker_and_industry <- data.frame("Ticker" = double(), 
                                  "Industry" = double())

for (val in seq(from=1, to=length(ticker_and_industry_list), by=2)) {
  
  row <- c(as.character(ticker_and_industry_list[val]), ticker_and_industry_list[val+1])
  ticker_and_industry[nrow(ticker_and_industry)+1,] <- row }

#ticker_and_industry

```


Complex linear regression model using data from the S&P 500 for both train and test sets. 
```{r linear Regression Analysis}

clean_key_stats <- read.csv("clean_key_stats.csv", sep=",", stringsAsFactors = FALSE, check.names = FALSE)

clean_key_stats <- dplyr::select(clean_key_stats, -c("5 Year Average Dividend Yield 4", "Forward Annual Dividend Rate 4",  "Forward Annual Dividend Yield 4",  "Last Split Factor (new per old) 2", "Quarterly Earnings Growth (yoy)", "Short % of Float (Mar 15, 2019) 4",  "Trailing Annual Dividend Rate 3",  "Trailing Annual Dividend Yield 3"))

clean_key_stats <- clean_key_stats[complete.cases(clean_key_stats),]
clean_key_stats[clean_key_stats == "N/A"]  <- NA
clean_key_stats <- na.omit(clean_key_stats)
clean_key_stats <- dplyr::select(clean_key_stats,  -c("ticker"))
clean_key_stats <- dplyr::select(clean_key_stats, -c("S&P500 52-Week Change 3"))

for (i in 2:ncol(clean_key_stats)) {
  for (j in 1:nrow(clean_key_stats)) {
    clean_key_stats[[i]][[j]] <- clean_number(clean_key_stats[[i]][[j]])
  }
}
#clean_key_stats

clean_key_stats <- clean_key_stats%>% mutate_all(funs(type.convert(as.double(.))))

norm_clean_key_stats <- clean_key_stats

for (i in 3:ncol(norm_clean_key_stats)) {
  max1 <- as.numeric(max(norm_clean_key_stats[[i]]))
  min1 <- as.numeric(min(norm_clean_key_stats[[i]]))
  for (j in 1:nrow(norm_clean_key_stats)) {
    norm_clean_key_stats[[i]][[j]] <- (as.numeric(norm_clean_key_stats[[i]][[j]]) - min1) / (max1 - min1)
  }
}


#clean_key_stats
sample_size <- floor(0.70 * nrow(norm_clean_key_stats))

classes <- sapply(norm_clean_key_stats, class)
set.seed(120)

new_set <- sample(seq_len(nrow(norm_clean_key_stats)), size = sample_size)
non_norm_train <- clean_key_stats[new_set, ]
initial_test <- norm_clean_key_stats[-new_set, ]
initial_train <- norm_clean_key_stats[new_set, ]

```


Complex model cont. Removes the least significant variables one at a time to ensure that only the most significant variables are considered and are not removed prematurely.
```{r Removing Insignificant Variables} 
updated_train <- initial_train
updated_test <- initial_test
train_control <- trainControl(method="repeatedcv", number=5, repeats=3)
has_insignificant_vars <- TRUE
while(has_insignificant_vars) {
  test_model <- caret::train(price~., data=updated_train, trControl=train_control, method="lm")
  
  prob <- data.frame(summary(test_model)$coefficients[,"Pr(>|t|)"])
  prob <- prob[1:nrow(prob),]
  prob <- data.frame(prob)
  
  variables <- summary(test_model)$coefficients[,0]
  
  variables <- as.data.frame(row.names(variables))
  
  prob <- cbind(variables, prob)
  
  names(prob) <- c("Variable", "Probability")
  prob
  prob <- data.frame(prob[2:nrow(prob),])
  max_p <- max(prob[,"Probability"]) 
  if (max_p <= 0.05) {
    has_insignificant_vars <- FALSE
  }
  else {
    insig_var <- prob[prob[,"Probability"] == max_p,1]
    updated_train <- dplyr::select(updated_train, -c(insig_var))
    updated_test <- dplyr::select(updated_test, -c(insig_var))
  }
}

#updated_train


# train the model
complex_model <- caret::train(price~., data=updated_train, trControl=train_control, method="lm")
summary(complex_model)
prediction <- complex_model %>% predict(updated_test)
#prediction
```

The simple model takes the most significant variables in the initial screenining of the model. 
```{r Simple Model}
library(caret)
library(tidyverse)
library(dplyr)

clean_key_stats <- read.csv("clean_key_stats3.csv", sep=",", stringsAsFactors = FALSE, check.names = FALSE)
clean_key_stats <- clean_key_stats[complete.cases(clean_key_stats),]
clean_key_stats <- na.omit(clean_key_stats)
clean_key_stats[clean_key_stats == "N/A"]  <- NA
clean_key_stats <- na.omit(clean_key_stats)

for (i in 2:ncol(clean_key_stats)) {
  for (j in 1:nrow(clean_key_stats)) {
    clean_key_stats[[i]][[j]] <- clean_number(clean_key_stats[[i]][[j]])
  }
}

clean_key_stats$`200-Day Moving Average 3` <- as.numeric(as.character(clean_key_stats$`200-Day Moving Average 3`))
clean_key_stats$`50-Day Moving Average 3` <- as.numeric(as.character(clean_key_stats$`50-Day Moving Average 3`))
clean_key_stats$`52 Week High 3` <- as.numeric(as.character(clean_key_stats$`52 Week High 3`))
clean_key_stats$`52 Week Low 3` <- as.numeric(as.character(clean_key_stats$`52 Week Low 3`))
clean_key_stats$`Total Debt/Equity (mrq)` <- as.numeric(as.character(clean_key_stats$`Total Debt/Equity (mrq)`))

for (i in 3:ncol(clean_key_stats)) {
  max1 <- as.numeric(max(clean_key_stats[[i]]))
  min1 <- as.numeric(min(clean_key_stats[[i]]))
  for (j in 1:nrow(clean_key_stats)) {
    clean_key_stats[[i]][[j]] <- (as.numeric(clean_key_stats[[i]][[j]]) - min1) / (max1 - min1)
  }
}
clean_key_stats <- dplyr::select(clean_key_stats, -c("S&P500 52-Week Change 3"))

clean_key_stats <- dplyr::select(clean_key_stats,  -c("ticker"))


sample_size <- floor(0.70 * nrow(clean_key_stats))

classes <- sapply(clean_key_stats, class)
set.seed(100)

new_set <- sample(seq_len(nrow(clean_key_stats)), size = sample_size)

key_stats_significant <- dplyr::select(clean_key_stats, c("price", "200-Day Moving Average 3", "50-Day Moving Average 3", "52 Week High 3", "52 Week Low 3", "52-Week Change 3", "Book Value Per Share (mrq)", "Diluted EPS (ttm)", "Enterprise Value/EBITDA 6"))

train <- key_stats_significant[new_set, ]
test <- key_stats_significant[-new_set, ]

simple_model <- lm(price ~., data = train)


```


Predicts a value using the complex model. If that value cannot be predicted due to missing data, the function will use the simple model to predict the value. 
```{r prediction}

value <- function(x) {
  
  key_stats <- data.frame(matrix(ncol=0, nrow=0))
  
  stats_url <- read_html(paste(paste(paste("https://finance.yahoo.com/quote/", x, sep=""), "/key-statistics?p=", sep = ""), x, sep =""))
  
  stats <- html_nodes(stats_url, "td") 
  clean_stats <- html_text(stats)
  
  price_url <- read_html(paste("https://www.nasdaq.com/symbol/", x, sep=""))
  price <- clean_number(html_text(html_nodes(price_url, "#qwidget_lastsale")))
  
  stats_df <- data.frame("Stat" = double(), "Value" = double())
  
  for (tick in seq(1, length(clean_stats), 2)) {
    row <- c(as.character(clean_stats[tick]), clean_value(clean_stats[tick+1]))
    stats_df[nrow(stats_df)+1,] <- row  
  }
  
  stats_df <- stats_df[stats_df$Stat != "Fiscal Year Ends ",]
  stats_df <- stats_df[stats_df$Stat != "Most Recent Quarter (mrq)",]
  stats_df <- stats_df[stats_df$Stat != "Dividend Date 3",]
  stats_df <- stats_df[stats_df$Stat != "Ex-Dividend Date 4",]
  stats_df <- stats_df[stats_df$Stat != "Last Split Date 3",]
  
  spread_df <- spread(stats_df, "Stat", "Value")
  if (nrow(key_stats) > 0) { 
    if (same_names(spread_df, key_stats)) {
      new_stats_df <- cbind(ticker = x, cbind(price, spread_df))
      key_stats <- rbind(key_stats, new_stats_df)
    }
  } else {
    new_stats_df <- cbind(ticker = x, cbind(price, spread_df))
    key_stats <- rbind(key_stats, new_stats_df)
  }
  
  test_value <- key_stats
  for (val in 2:ncol(test_value)) {
    
    for (row in 1:nrow(test_value)) {
      test_value[[val]][[row]] <- clean_number(as.character(test_value[[val]][[row]]))
    }
  }
  
  
  test_value <- test_value[colnames(updated_train)]
  test_value <- test_value%>% mutate_all(funs(type.convert(as.double(.))))
  
  normalized_data <- rbind(non_norm_train[colnames(updated_train)]%>% mutate_all(funs(type.convert(as.double(.)))), test_value)
  
  for (i in 2:ncol(normalized_data)) {
    max1 <- as.numeric(max(normalized_data[[i]]))
    min1 <- as.numeric(min(normalized_data[[i]]))
    for (j in 1:nrow(normalized_data)) {
      normalized_data[[i]][[j]] <- (as.numeric(normalized_data[[i]][[j]]) - min1) / (max1 - min1)
    }
  }
  
  new_val <- normalized_data[nrow(normalized_data),]
  
  
  c_predict <- predict(complex_model, new_val)
  
  if (length(c_predict) == 0) {
    
    c_predict <- predict(simple_model, new_val)
  }
  c_predict
}

```


Tests results to ensure validity. 
```{r test}
#returns 183.41 compared to value at approx 180 (would issue a buy recommendation)
value("fb")

#returns 199.45 compared to value at approx 199 (would issue a sell/hold recommendation)
value("aapl")

#returns 32.54 compared to value at approx 35 (would issue a sell/hold recommendation)
value("twtr")

#returns 286.68 compared to value at approx 267 (would issue a sell/hold recommendation)
value("tsla")

```


Serves as a dictionary in conjunction with the S&P ticker list for RShiny queries.
```{r russell 1000}

russell_ticker_url <- read_html("https://en.wikipedia.org/wiki/Russell_1000_Index")
russell_ticker_list <- html_text(html_nodes(russell_ticker_url, "td+ td"))

russell_ticker_list <- russell_ticker_list[51:length(russell_ticker_list)]
for (val in 1:length(russell_ticker_list)) {
  russell_ticker_list[val] <- substr(russell_ticker_list[val], 0, nchar(russell_ticker_list[val])-1)
}


russell_df <- data.frame("Ticker" = double())
for (val in russell_ticker_list) {
  russell_df[nrow(russell_df)+1,] <- c(val)
}

```


Compares our predicted value against the current market price to provide the user awith a recommendation.
```{r}

compare_vals <- function(x) {
  
  price_url <- read_html(paste("https://www.nasdaq.com/symbol/", x, sep=""))
  price <- clean_number(html_text(html_nodes(price_url, "#qwidget_lastsale")))
  recommendation <- if (value(x) > price) {
    recommendation <- "Buy"
  }
  else {
    recommendation <- "This Stock is Currently Over-Valued"
  }
  
  paste(paste(paste(paste("Estimated Current Value: ", round(value(x), 2), sep =""), paste(" | Current Price: ", price, sep =""), sep=""), " | Recommendation: ", sep=""), recommendation, sep="")
}

```




